---
title: "TCE_HW4"
author: "KiseokUchicago"
date: "2021-04-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

## Homework4 (3b, 4a, 4b)
## Coding assignment for ECEV 43900 Theoretical Community Ecology
Professor: **Stefano Allesina** \
Student: **Kiseok Lee**  

## Problem 3b

Consider a five-species system with symmetric, stable B (with all positive coefficients) and positive s, yielding the feasible equilibrium x*.\

I did not understand what stable B means. Here I used the D-stability where B + BT is negative definite.

(1) Finding random parameters that satisfies:
- B is symmetric, positive elements, stable
- s is positive (B-1s=x*) Bx = s
- x* (nontrivial fixed point exists = positive)

```{r}
library(matrixcalc)

# five species system
n = 5

# randomizing s
s = runif(n, 0.001, 1)
# randomizing B
B = matrix(runif(n*n, 0.001, 1),n, n)
# symmetric
B[lower.tri(B)] = 0
B <- B + t(B)
diag(B) <- runif(n, 0.001, 1)
isSymmetric(B)
# x* is positive
all(solve(B, s) > 0) == T
# B is stable <=> real part of eigenvalues are all negative
# function
rightmost <- function(M){
  eig <- eigen(M, only.values = T)$values
  # print(Re(eig)) # print all real part of eigenvalues
  return(max(Re(eig))) # the right most eigenvalue
} 
rightmost(B) < 0

# finding
# while(!((all(solve(B, s) > 0) == T) & (rightmost(B) < 0))){
#   # randomizing s
#   # s = runif(n, 0.001, 1)
#   # randomizing B
#   B = matrix(runif(n*n, 0.001, 1),n, n)
#   # symmetric
#   B[lower.tri(B)] = 0
#   B <- B + t(B)
#   diag(B) <- runif(n, 0.001, 1)
#   print(B)
# }
```

I couldn't get the while function to find B and s that satisfied all condition.\
I loosened the B stable condition.
```{r}
while(!(all(solve(B, s) > 0) == T)){
  # randomizing s
  s = runif(n, 0.001, 1)
  # randomizing B
  B = matrix(runif(n*n, 0.001, 1),n, n)
  # symmetric
  B[lower.tri(B)] = 0
  B <- B + t(B)
  diag(B) <- runif(n, 0.001, 1)
  # print(B)
}

# Found B
B
s
```


(2) For each possible subset of species in the pool, (i.e., for all combinations ranging from a single species [5 cases], to two species [10 cases], to all species together [1 case]), compute the corresponding equilibrium. Is it feasible? Is it stable?

```{r}

# for all species
all(solve(B, s) > 0) == T # feasible
rightmost(B) < 0 # but not stable (because of the loosened condition to find B)

# for 4 species combination
c4 <- combn(1:5, 4)
B
for (i in 1:dim(c4)[2]){
  # new species combination matrix
  B_c4 <- B[c4[,i],c4[,i]]
  s4 <- s[c4[,i]]
  print(paste0('case',i))
  print(B_c4)
  print(paste0("Feasibility: ",all(solve(B_c4, s4) > 0) == T))
  print(paste0("Stability: ",rightmost(B_c4) < 0))
}

# for 3 species combination
c3 <- combn(1:5, 3)
B
for (i in 1:dim(c3)[2]){
  # new species combination matrix
  B_c3 <- B[c3[,i],c3[,i]]
  s3 <- s[c3[,i]]
  print(paste0('case',i))
  print(B_c3)
  print(paste0("Feasibility: ",all(solve(B_c3, s3) > 0) == T))
  print(paste0("Stability: ",rightmost(B_c3) < 0))
}

# for 2 species combination
c2 <- combn(1:5, 2)
B
for (i in 1:dim(c2)[2]){
  # new species combination matrix
  B_c2 <- B[c2[,i],c2[,i]]
  s2 <- s[c2[,i]]
  print(paste0('case',i))
  print(B_c2)
  print(paste0("Feasibility: ",all(solve(B_c2, s2) > 0) == T))
  print(paste0("Stability: ",rightmost(B_c2) < 0))
}

# for 1 species combination
c1 <- combn(1:5, 1)
B
for (i in 1:dim(c1)[2]){
  # new species combination matrix
  B_c1 <- B[c1[,i],c1[,i]]
  s1 <- s[c1[,i]]
  print(paste0('case',i))
  print(B_c1)
  print(paste0("Feasibility: ",all(solve(B_c1, s1) > 0) == T))
  print(paste0("Stability: ",rightmost(B_c1) < 0))
}

```

(3) Take two subset of species such that a) both are feasible and stable; b) subset 1 contains subset 2 (i.e., all species in 2 are in 1, but not the converse); c) the value of V(x*) for subset 1 is larger than that for subset 2. Try invading subset 2 with the species by introducing at the equilibrium of subset 2 the species that are in subset 1 but not in 2—starting all of them at low density. What happens?

```{r}

```


## Problem 4a

(1) Write code to compute the real part for the “rightmost” eigenvalue of a random matrix (Note: computing eigenvalues is fairly expensive in terms of computing time. Use eigen(M, only.values = TRUE)$values to speed up calculations).
```{r}
# define matrix M and A
# M = A * x
n = 5
M = matrix(runif(n*n, 0, 1),n, n)

# function
rightmost <- function(M){
  eig <- eigen(M, only.values = T)$values
  # print(Re(eig)) # print all real part of eigenvalues
  return(max(Re(eig))) # the right most eigenvalue
} 

rightmost(M)
```

(2) Write code to build matrices like those studied by May (nonzero elements sampled from a normal distribution)

May's matrix: Mij = 0 with probability (1-C); with probability C we draw Mij from a distribution with mean zero and variance σ^2. \  
C is the proportion of realized connections, termed the “connectance” of the system.\
the diagonal elements are set to −d, modeling self-regulation.

```{r}
n = 5
sigma = 1
C = 0.3   # connectance
d = 10

# May matrix generating function
may_matrix <- function(sigma, C, n, d){
  M = matrix(rnorm(n*n, mean=0, sd=sigma),n, n)
  # then make elements 0 with probability of 1-C
  M_prob <- matrix(rbinom(n*n, size=1, prob=C),n, n)
  # apply probability
  May_M <- M * M_prob
  diag(May_M) <- -d 
  return(May_M)
}

may_matrix(sigma, C, n, d)

```

(3) Set d=10 and choose parameters n, C, and σ^2 such that you are close to the critical point (make sure n and C are large enough, for example nC > 10). Draw 1000 random matrices and compute the probability drawing a stable matrix.

```{r}
d = 10
n = 300
C = 0.4
sigma=0.9

# stability condition
sqrt(n*C*(sigma^2)) # 9.859006
d # 10
sqrt(n*C*(sigma^2)) < d # True

# generate 1000 random matrices
num_random = 1000
stable_list = rep(-1,num_random) # vector that stores information of stable (true=1) vs unstable (false=0)

for (i in 1:num_random){
  m <- may_matrix(sigma, C, n, d)
  # print(rightmost(m))
  stable_list[i] <- rightmost(m)<0
}

# probability drawing a stable matrix
sum(stable_list)/num_random

```

(4) Vary n, C, and sigma^2 in turn, making them cross the critical point. Draw a graph where the probability of stability is on the y axis, the x axis measures sqrt(nCsigma^2). The graph should look like the one reported below:

```{r}
# function for calculating stability
prob_stability <- function(num_random=1000, sigma, C, n, d=10){
  stable_list = rep(-1,num_random) # vector that stores information of stable (true=1) vs unstable (false=0)
  for (i in 1:num_random){
    m <- may_matrix(sigma, C, n, d)
    # print(rightmost(m))
    stable_list[i] <- rightmost(m)<0
  }
  return(sum(stable_list)/num_random)
}

# creating dataframe for changing values
# (1) changing C from 0.3 to 0.5. n = 300, sigma=0.9
vec_C <- seq(0.25, 0.55, length.out=16)
n = 300; sigma=0.9
# x axis
sqrt(n*vec_C*(sigma^2))
# y axis
vec_prob1 <- rep(-1,length(vec_C))
for (i in 1:length(vec_C)){
  vec_prob1[i] <- prob_stability(num_random=100, sigma=0.9, C=vec_C[i], n = 300, d=10)
}

# (2) Changing sigma from 0.7 to 1.1 & n = 300; C = 0.4
vec_sigma <- seq(0.7, 1.1, length.out=16)
n = 300; C = 0.4
# x axis
sqrt(n*C*(vec_sigma^2))
# y axis
vec_prob2 <- rep(-1,length(vec_sigma))
for (i in 1:length(vec_sigma)){
  vec_prob2[i] <- prob_stability(num_random=100, sigma=vec_sigma[i], C=0.4, n = 300, d=10)
}

# (3) changing n from 180 to 430 & C = 0.4; sigma=0.9
vec_n <- seq(180, 430, length.out=16)
C = 0.4; sigma=0.9
# x axis
sqrt(vec_n*C*(sigma^2))
# y axis
vec_prob3 <- rep(-1,length(vec_n))
for (i in 1:length(vec_n)){
  vec_prob3[i] <- prob_stability(num_random=100, sigma=0.9, C=0.4, n = vec_n[i], d=10)
}

## bind the results into single dataframe
df1 <- data.frame(x=sqrt(n*vec_C*(sigma^2)),y=vec_prob1,varying="C")
df2 <- data.frame(x=sqrt(n*C*(vec_sigma^2)),y=vec_prob2,varying="sigma")
df3 <- data.frame(x=sqrt(vec_n*C*(sigma^2)),y=vec_prob3,varying="n")
df_merge <- rbind(df1, df2, df3)

library(ggplot2)
# plot
ggplot(df_merge,aes(x = x,y = y)) + geom_point(aes(shape=varying), size=3, color='maroon') + theme_bw() +
  labs(title="Probability of matrix being stable depending on C, n, and sigma \n", x="\n sqrt(n*C*(sigma^2))",y="probability of having stable community matrix \n")

```

## Problem 4b

Study the spectrum of the community matrix of a competitive community in which species are divided into two groups (e.g., benthic, pelagic), and interactions are as follow:

Community matrix conditions:
1. Mii (diagonal) = -5 \
2. Mij when species i and j is in the same group, Mij ~ N(mu1, sigma^2)\
3. Mij when species i and j is in the different group, Mij ~ N(mu2, sigma^2)\
4. There are only two groups and the number of species in each group is the same.

(1) Now we are constructing matrix with the conditions above:
```{r}
# parameters setting
n = 10
sigma = 1/2
mu1 = -1   # connectance
mu2 = -1/2

# Matrix generating function
# let's say everytime we cluster the species in the same group together in the matrix.
# so that we can make structured matrix.

twogroup_matrix <- function(sigma=0.5, n, mu1=-1,mu2=-0.5){
  M = matrix(rep(0,n*n),n,n)
  for (i in 1:n){
    for (j in 1:n){
      if(i != j){
        # print(paste0(i,',',j))
        # group 1 is species {1, 2, ..., n/2}
        # group 2 is species {(n/2)+1, ..., n}
        if ((i<= n/2 & j<= n/2) | (i> n/2 & j> n/2)){ # for i and j in the same group
          M[i,j] <- rnorm(1, mean=mu1, sd=sigma)
          # print(M[i,j]) # testing
        } else{ # # for i and j in the different group
          M[i,j] <- rnorm(1, mean=mu2, sd=sigma)
          # print(M[i,j]) # testing
        }
      }
    }
  }
  diag(M) <- -5 
  return(M)
}

twogroup_matrix(sigma=0.5, n=10, mu1=-1,mu2=-0.5)

```

(2) Analyze the eigenvalue space:

```{r}

eig <- eigen(twogroup_matrix(sigma=0.5, n=1000, mu1=-1,mu2=-0.5), only.values = T)$values
df_eig <- data.frame(Real_lambda = Re(eig), Im_lambda = Im(eig))

ggplot(df_eig,aes(x = Real_lambda,y = Im_lambda)) + geom_point(size=3, color='maroon') + theme_bw() +
  scale_x_continuous(limits = c(min(Re(eig)), abs(min(Re(eig)))))+
  labs(title="Eigenvalue distribution \n", x="\n Real part of eigenvalue",y="Imaginary part of eigenvalue \n")
```

(3) Do with multiple random matrices:
```{r}
n_random = 50

df_eig <- data.frame(Real_lambda = NULL, Im_lambda = NULL, num_random=NULL)
for (i in 1:n_random){
  # 100 species commuity matrix's eigen value
  eig <- eigen(twogroup_matrix(sigma=0.5, n=100, mu1=-1,mu2=-0.5), only.values = T)$values
  df_eig_add <- data.frame(Real_lambda = Re(eig), Im_lambda = Im(eig), num_random=i)
  df_eig <- rbind(df_eig, df_eig_add)
}

ggplot(df_eig,aes(x = Real_lambda,y = Im_lambda)) + geom_point(size=3, aes(color=num_random)) + theme_bw() +
  labs(title="Eigenvalue distribution \n", x="\n Real part of eigenvalue",y="Imaginary part of eigenvalue \n")
```

(4) The left most eigenvalue value changes with change of n
```{r}
df_leftmost <- data.frame(n_species = NULL, leftmost_re_eig =NULL)
# from 10 species to 1000 species (only even number of species can devide into 2 groups)
for (n_species in seq(10,500,by=10)){
  # 10 randomization
  n_rand = 10
  min_re_eig = rep(0,n_rand)
  for (i in (1:n_rand)){
    eig_val <- eigen(twogroup_matrix(sigma=0.5, n=n_species, mu1=-1,mu2=-0.5), only.values = T)$values
    min_re_eig[i] <- min(Re(eig_val))
  }
  df_leftmost_add <- data.frame(n_species = rep(n_species,n_rand), leftmost_re_eig = min_re_eig)
  df_leftmost <- rbind(df_leftmost, df_leftmost_add)
}

# plot
ggplot(df_leftmost,aes(x = n_species,y = leftmost_re_eig)) + geom_point(size=1, shape=1) + theme_bw() +
  geom_smooth(method='lm', formula= y~x, size=0.1, color = 'red')+
  labs(title="Left most eigenvalue's real part changes with matrix size \n", x="\n Number of species (community matrix size)",y="Left most real part of eigenvaluee \n")

```








