---
title: "TCE_HW7"
author: "KiseokUchicago"
date: "2021-06-02"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

## Homework7 (Game theory)
## Coding assignment for ECEV 43900 Theoretical Community Ecology
Professor: **Stefano Allesina** (TA: **Zach Miller**) \
Student: **Kiseok Lee**  

```{r}
# libraries
library(igraph)
library(tidygraph)
library(ggraph)
library(bipartite)
library(dplyr)
library(ggplot2)
library(ggrepel)
library(RColorBrewer)
library(vegan)
library(tidyverse)
library(magrittr)
library(reshape2)
library(gtools)
library(devtools)
library(ggpubr)
library(MASS)
library(deSolve)

## theme for ggplot
mytheme <- theme_bw() + 
  theme(plot.title = element_text(size = 19,hjust = 0.5, family="serif")) + 
  theme(axis.title.x = element_text(size = 17,hjust = 0.5, family="serif")) + 
  theme(axis.title.y = element_text(size = 17,hjust = 0.5, family="serif")) + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust=0.3,size=13, family="serif"))+
  theme(axis.text.y = element_text(size=10, family="serif"))+
  theme(panel.grid.major = element_blank()) +
  theme(panel.grid.minor = element_blank(),panel.background=element_blank(),panel.border=element_blank(),plot.background=element_blank()) +
  theme(axis.ticks = element_line(size = 1.1))
  
mytheme_2d <- theme_bw() + 
  theme(plot.title = element_text(size = 19,hjust = 0.5, family="serif")) + 
  theme(axis.title.x = element_text(size = 17,hjust = 0.5, family="serif")) + 
  theme(axis.title.y = element_text(size = 17,hjust = 0.5, family="serif")) + 
  theme(axis.text.x = element_text(hjust = 0.5, vjust=0.3,size=13, family="serif"))+
  theme(axis.text.y = element_text(size=13, family="serif"))+
  theme(panel.grid.major = element_blank()) +
  theme(panel.grid.minor = element_blank(),panel.background=element_blank(),plot.background=element_blank()) +
  theme(axis.ticks = element_line(size = 1.1))


# color collection
my_color_collection <- c(
  "#CBD588", "#5F7FC7", "orange", "#AD6F3B", "#673770", 
  "#D14285", "#652926", "#C84248", "#8569D5", "#5E738F",
  "#D1A33D", "#8A7C64", "#599861","#616163", "#FFCDB2",
  "#6D9F71", "#242F40",
   "#CCA43B", "#F92A82", "#ED7B84", "#7EB77F", 
  "#DEC4A1", "#E5D1D0", '#0E8482', '#C9DAEA', '#337357', 
  '#95C623', '#E55812', '#04471C', '#F2D7EE', '#D3BCC0', 
  '#A5668B', '#69306D', '#0E103D', '#1A535C', '#4ECDC4', 
  '#F7FFF7', '#FF6B6B', '#FFE66D', '#6699CC', '#FFF275', 
  '#FF8C42', '#FF3C38', '#A23E48', '#000000', '#CF5C36', 
  '#EEE5E9', '#7C7C7C', '#EFC88B', '#2E5266', '#6E8898', 
  '#9FB1BC', '#D3D0CB', '#E2C044', '#5BC0EB', '#FDE74C', 
  '#9BC53D', '#E55934', '#FA7921', "#CD9BCD", "#508578", "#DA5724")

```

## 7a
Analyze the replicator equation when the matrix A.
for the cases in which a) α>1, b) α<1. Write code to simulate the system, and prove stability/instability depending on the value of α.

Set up functions for replicator dynamics
```{r}
# define the differential equation
RE <-function(t, x, parameters){
  with(as.list(c(x, parameters)), {
  x[x < 10^-8] <- 0 # prevent numerical problems
  x <- x / sum(x) # keep on simplex
  dxdt <- x * (A %*% x - sum(x * A %*% x))
  list(dxdt)
  })
}

# general function to integrate RE and plot
integrate_RE_and_plot <- function(alpha, x0, maxtime = 200, steptime = 0.05){
  times <- seq(0, maxtime, by = steptime)
  A = matrix(c(0, 1, -alpha,
             -alpha, 0, 1,
             1, -alpha, 0), nrow=3, ncol=3)
  parameters <- list(A = A)
  # solve numerically
  # Initial conditions
  set.seed(1)
  x0 <- runif(3)
  x0 <- x0 / sum(x0)
  out <- ode(y = x0, times = times, func = RE, parms = parameters, method = "ode45")
  # plot and make into tidy form
  out <- as.data.frame(out)
  colnames(out) <- c("time", paste("sp", 1:(ncol(out) -1), sep = "_"))
  out <- as_tibble(out) %>% gather(species, density, -time)
  
  p <- ggplot(data = out) + aes(x = time, y = density, colour = species) +
    geom_line() +
    scale_x_continuous(expression(t)) +
    scale_y_continuous(expression(x(t))) +
    mytheme_2d
  return(p)
}


```

Plot and see.
only 1 species exist.
```{r}
## (1) When alpha is positive
integrate_RE_and_plot(alpha=3, x0)

## (2) When alpha is negative
integrate_RE_and_plot(alpha=-3, x0)

## (3) some range of alphas
alpha_vec <- seq(-4, 4, by=1)
length(alpha_vec)

p1 <- integrate_RE_and_plot(alpha=alpha_vec[1], x0) 
p2 <- integrate_RE_and_plot(alpha=alpha_vec[2], x0) 
p3 <- integrate_RE_and_plot(alpha=alpha_vec[3], x0)
p4 <- integrate_RE_and_plot(alpha=alpha_vec[4], x0) 
p5 <- integrate_RE_and_plot(alpha=alpha_vec[5], x0) 
p6 <- integrate_RE_and_plot(alpha=alpha_vec[6], x0) 
p7 <- integrate_RE_and_plot(alpha=alpha_vec[7], x0) 
p8 <- integrate_RE_and_plot(alpha=alpha_vec[8], x0) 
p9 <- integrate_RE_and_plot(alpha=alpha_vec[9], x0)

ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9, # + rremove("x.text") 
          labels = alpha_vec,
          ncol = 3, nrow = 3)
```

Stability analysis

dx/dt = D(x)Ax
Jacobian matrix = D(x*)A
```{r}
# feasible fixed points
alpha <- -2 
A = matrix(c(0, 1, -alpha,
           -alpha, 0, 1,
           1, -alpha, 0), nrow=3, ncol=3)
# A %*% x = sum(x * A %*% x)

```
Can't find the feasible equilibrium...?

## 7b. 
As seen above, generally zero-sum matrix games will lead to an odd number of species coexisting.

Can you build a matrix leading to four species/strategies coexisting?
Now simulate the dynamics several times, starting from different initial conditions—do species cycle around the same equilibrium? Why?

```{r}
# Try A with 4 species
# general function to integrate RE and plot
integrate_RE_and_plot <- function(alpha, x0, maxtime = 200, steptime = 0.05){
  times <- seq(0, maxtime, by = steptime)
  A = matrix(c(0, 0, 1, -alpha,
             -alpha, 0, 0, 1,
             1, -alpha, 0, 0,
             0, 0, 1, -alpha), nrow=4, ncol=4)
  parameters <- list(A = A)
  # solve numerically
  # Initial conditions
  set.seed(1)
  x0 <- runif(4)
  x0 <- x0 / sum(x0)
  out <- ode(y = x0, times = times, func = RE, parms = parameters, method = "ode45")
  # plot and make into tidy form
  out <- as.data.frame(out)
  colnames(out) <- c("time", paste("sp", 1:(ncol(out) -1), sep = "_"))
  out <- as_tibble(out) %>% gather(species, density, -time)
  
  p <- ggplot(data = out) + aes(x = time, y = density, colour = species) +
    geom_line() +
    scale_x_continuous(expression(t)) +
    scale_y_continuous(expression(x(t))) +
    mytheme_2d
  return(p)
}

##  some range of alphas
alpha_vec <- seq(-4, 4, by=1)
length(alpha_vec)

p1 <- integrate_RE_and_plot(alpha=alpha_vec[1], x0) 
p2 <- integrate_RE_and_plot(alpha=alpha_vec[2], x0) 
p3 <- integrate_RE_and_plot(alpha=alpha_vec[3], x0)
p4 <- integrate_RE_and_plot(alpha=alpha_vec[4], x0) 
p5 <- integrate_RE_and_plot(alpha=alpha_vec[5], x0) 
p6 <- integrate_RE_and_plot(alpha=alpha_vec[6], x0) 
p7 <- integrate_RE_and_plot(alpha=alpha_vec[7], x0) 
p8 <- integrate_RE_and_plot(alpha=alpha_vec[8], x0) 
p9 <- integrate_RE_and_plot(alpha=alpha_vec[9], x0)

ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,
          labels = alpha_vec,
          ncol = 3, nrow = 3)

```

Try expanding the matrix A size

```{r}
# Try A with 5 species
# general function to integrate RE and plot
integrate_RE_and_plot <- function(alpha, x0, maxtime = 200, steptime = 0.05){
  times <- seq(0, maxtime, by = steptime)
  A = matrix(c(0, 0, 0, 1, -alpha,
             -alpha, 0, 0, 0, 1,
             1, -alpha, 0, 0, 0,
             0, 1, -alpha, 0, 0,
             0, 0, 1, -alpha, 0,
             0, 0, 0, 1, -alpha), nrow=5, ncol=5)
  parameters <- list(A = A)
  # solve numerically
  # Initial conditions
  set.seed(1)
  x0 <- runif(5)
  x0 <- x0 / sum(x0)
  out <- ode(y = x0, times = times, func = RE, parms = parameters, method = "ode45")
  # plot and make into tidy form
  out <- as.data.frame(out)
  colnames(out) <- c("time", paste("sp", 1:(ncol(out) -1), sep = "_"))
  out <- as_tibble(out) %>% gather(species, density, -time)
  
  p <- ggplot(data = out) + aes(x = time, y = density, colour = species) +
    geom_line() +
    scale_x_continuous(expression(t)) +
    scale_y_continuous(expression(x(t))) +
    mytheme_2d
  return(p)
}

##  some range of alphas
alpha_vec <- seq(-4, 4, by=1)
length(alpha_vec)

p1 <- integrate_RE_and_plot(alpha=alpha_vec[1], x0) 
p2 <- integrate_RE_and_plot(alpha=alpha_vec[2], x0) 
p3 <- integrate_RE_and_plot(alpha=alpha_vec[3], x0)
p4 <- integrate_RE_and_plot(alpha=alpha_vec[4], x0) 
p5 <- integrate_RE_and_plot(alpha=alpha_vec[5], x0) 
p6 <- integrate_RE_and_plot(alpha=alpha_vec[6], x0) 
p7 <- integrate_RE_and_plot(alpha=alpha_vec[7], x0) 
p8 <- integrate_RE_and_plot(alpha=alpha_vec[8], x0) 
p9 <- integrate_RE_and_plot(alpha=alpha_vec[9], x0)

ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,
          labels = alpha_vec,
          ncol = 3, nrow = 3)
```

With 8 species, 4 species are coexisting
```{r}
# Try A with 8 species
# general function to integrate RE and plot
integrate_RE_and_plot <- function(alpha, x0, maxtime = 200, steptime = 0.05){
  times <- seq(0, maxtime, by = steptime)
  A = matrix(c(0, 0, 0, 0, 0, 0, 1, -alpha,
             -alpha, 0, 0, 0, 0, 0, 0, 1,
               1, -alpha, 0, 0, 0, 0, 0, 0,
               0, 1, -alpha, 0, 0, 0, 0, 0,
               0, 0, 1, -alpha, 0, 0, 0, 0,
               0, 0, 0, 1, -alpha, 0, 0, 0,
               0, 0, 0, 0, 1, -alpha, 0, 0,
               0, 0, 0, 0, 0, 1, -alpha, 0), nrow=8, ncol=8)
  parameters <- list(A = A)
  # solve numerically
  # Initial conditions
  set.seed(1)
  x0 <- runif(8)
  x0 <- x0 / sum(x0)
  out <- ode(y = x0, times = times, func = RE, parms = parameters, method = "ode45")
  # plot and make into tidy form
  out <- as.data.frame(out)
  colnames(out) <- c("time", paste("sp", 1:(ncol(out) -1), sep = "_"))
  out <- as_tibble(out) %>% gather(species, density, -time)
  
  p <- ggplot(data = out) + aes(x = time, y = density, colour = species) +
    geom_line() +
    scale_x_continuous(expression(t)) +
    scale_y_continuous(expression(x(t))) +
    mytheme_2d
  return(p)
}

##  some range of alphas
alpha_vec <- seq(-4, 4, by=1)
length(alpha_vec)

p1 <- integrate_RE_and_plot(alpha=alpha_vec[1], x0) 
p2 <- integrate_RE_and_plot(alpha=alpha_vec[2], x0) 
p3 <- integrate_RE_and_plot(alpha=alpha_vec[3], x0)
p4 <- integrate_RE_and_plot(alpha=alpha_vec[4], x0) 
p5 <- integrate_RE_and_plot(alpha=alpha_vec[5], x0) 
p6 <- integrate_RE_and_plot(alpha=alpha_vec[6], x0) 
p7 <- integrate_RE_and_plot(alpha=alpha_vec[7], x0) 
p8 <- integrate_RE_and_plot(alpha=alpha_vec[8], x0) 
p9 <- integrate_RE_and_plot(alpha=alpha_vec[9], x0)

ggarrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,
          labels = alpha_vec,
          ncol = 3, nrow = 3)
```



