---
title: "TCE_HW1"
author: "KiseokUchicago"
date: "2021-04-08"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

## Homework1 (Models for single population)
## Coding assignment for ECEV 43900 Theoretical Community Ecology
Professor: **Stefano Allesina** \
Student: **Kiseok Lee**  

```{r}
# libraries
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
```

## Homework 1b

Draw the bifurcation diagram for logistic growth model with harvesting:

(1) write code to solve the initial value problem from a random initial condition (i.e., sampled independently from a uniform distribution between 0 and 2); document the possible outcomes.
(2) write code to plot the bifurcation diagram—for each value of η, mark stable and unstable equilibria.
(3) what is the critical η where a bifurcation occurs?

### (1) write code to solve the initial value problem from a random initial condition (i.e., sampled independently from a uniform distribution between 0 and 2); document the possible outcomes.

```{r}
library(deSolve)

# get the parameters for the model
parameters <- c(rho = 1, alpha = 1, eta = 0.5)
state <- c(x = runif(1,0,2))

# set up the ode
harvesting_ode <- function(t, state, parameters) {
  with(as.list(c(state, parameters)),{
  # rate of change
  dx <- rho*x*(1-alpha*x)-eta*x
  list(c(dx))
  }) # end with(as.list ...
}

times <- seq(0, 100, by = 0.01)
outc <- ode(state, times, harvesting_ode, parameters, method = "radau", atol = 1e-4, rtol = 1e-4)
plot(outc, xlab = "time", ylab = "-")

## plot when eta is 0
times <- seq(0, 30, by = 0.01)
plot(0,0, xlim=c(0,30), ylim=c(-0.05,2.05),type="n", xlab="t", ylab="X", main="x(0) randomly sampled in [0,2] when eta=0") 
for (k in 1:50) { # number of random sampling of x(0)
  parameters <- c(rho = 1, alpha = 1, eta = 0)
  state <- c(x = runif(1,0,2))
  out <- ode(state, times, harvesting_ode, parameters, method = "radau", atol = 1e-4, rtol = 1e-4)
  points(out, pch=".")
}

## plot when eta is 0.2
times <- seq(0, 30, by = 0.01)
plot(0,0, xlim=c(0,30), ylim=c(-0.05,2.05),type="n", xlab="t", ylab="X", main="x(0) randomly sampled in [0,2] when eta=0.2") 
for (k in 1:50) { # number of random sampling of x(0)
  parameters <- c(rho = 1, alpha = 1, eta = 0.2)
  state <- c(x = runif(1,0,2))
  out <- ode(state, times, harvesting_ode, parameters, method = "radau", atol = 1e-4, rtol = 1e-4)
  points(out, pch=".")
}

## plot when eta is 0.5
times <- seq(0, 30, by = 0.01)
plot(0,0, xlim=c(0,30), ylim=c(-0.05,2.05),type="n", xlab="t", ylab="X", main="x(0) randomly sampled in [0,2] when eta=0.5") 
for (k in 1:50) { # number of random sampling of x(0)
  parameters <- c(rho = 1, alpha = 1, eta = 0.5)
  state <- c(x = runif(1,0,2))
  out <- ode(state, times, harvesting_ode, parameters, method = "radau", atol = 1e-4, rtol = 1e-4)
  points(out, pch=".")
}

## plot when eta is 1
times <- seq(0, 30, by = 0.01)
plot(0,0, xlim=c(0,30), ylim=c(-0.05,2.05),type="n", xlab="t", ylab="X", main="x(0) randomly sampled in [0,2] when eta=1") 
for (k in 1:50) { # number of random sampling of x(0)
  parameters <- c(rho = 1, alpha = 1, eta = 1)
  state <- c(x = runif(1,0,2))
  out <- ode(state, times, harvesting_ode, parameters, method = "radau", atol = 1e-4, rtol = 1e-4)
  points(out, pch=".")
}

```
x is converging to (rho - eta) / alpha * rho


### (2) write code to plot the bifurcation diagram—for each value of η, mark stable and unstable equilibria.

```{r}
## Plot the Bifurcation Diagram 
plot(0,0, xlim=c(-0.05,1.55), ylim=c(-0.05,2.05),type="n", xlab="eta", ylab="X", main="Bifurcation plot with eta") 
times <- seq(0, 100, by = 0.01)
for (Eta in seq(0,1.5,0.05)) { # These are the initial and final values for eta
  parameters <- c(rho = 1, alpha = 1, eta = Eta)
  state <- c(x = 1)
  # state <- c(x = runif(1,0,2))
  out <- ode(state, times, harvesting_ode, parameters, method = "radau", atol = 1e-4, rtol = 1e-4)
  points(Eta,out[dim(out)[1],2])
}
# indicating stable fixed points
x <- 0:1;y <- 1:0
clip(0,1,0,1)
abline(lm(y~x))
clip(0,1,0,0)
abline(h=0,lty = 2)
clip(1,1.5,0,0)
abline(h=0,lty = 1)

```

### (3) what is the critical η where a bifurcation occurs?
η = 1


