---
title: "TCE_HW5"
author: "KiseokUchicago"
date: "2021-05-13"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=9,
                      error=TRUE, echo=TRUE, eval=TRUE, warning=FALSE, message=FALSE)
```

## Homework5
## Coding assignment for ECEV 43900 Theoretical Community Ecology
Professor: **Stefano Allesina** (TA: **Zach Miller**) \
Student: **Kiseok Lee**  

Build the assembly graph for a GLV model in which the growth rates of all species are the same, and interactions are given by phylogenetic relatedness.

In particular, consider an ultrametric phylogenetic tree T, and the “variance-covariance” matrix induced by the tree, Σ. Because the tree is ultrametric, then Σii=1 for all i. The off-diagonal elements Σij express the proportion of shared ancestry (and as such 1≥Σij≥0).\
Take the GLV model:dxdt=D(x)(1−Σx)

Notice that Σ is positive (semi-)definite by construction. Now build the assembly graph: does it have any interesting property?

Let's start with constructing binary matrix that represents a tree
```{r}
# We are going to use a bottom up pairing approach to construct a tree
# the number of branches are 2n-1 (number of species = n)

# Create vector for each species (the most bottom branch)
n = 5
(sampling_matrix <- diag(n))

## (Step 1) From these species' most bottom branches, choose two species/clusters to group them. We are going to hierarchically group it, until we group every species into one group (this vector will be [1, 1, 1, 1, ..., 1])
set.seed(1)
(random_2_cols <- sample(dim(sampling_matrix)[2], 2, replace = FALSE, prob = NULL))

# (Step 2) put sampled columns into "accumulative_matrix"
(accumulative_matrix <- sampling_matrix[,random_2_cols])

# (Step 3) sum up picked two species/clusters and add it to the "sampling_matrix"
(sampling_matrix <- cbind(sampling_matrix, rowSums(sampling_matrix[,random_2_cols])))

# (Step 4) remove the randomly picked 2 columns from the "sampling_matrix"
(sampling_matrix <- sampling_matrix[,-random_2_cols, drop = FALSE]) # make the matrix not collapse into vector

# repeat this until rowSums(sampling_matrix[,random_2_cols]) is all 1
while (!all(sampling_matrix ==1)){
  ## (Step 1) From these species, choose two species/clusters to group them.
  random_2_cols <- sample(dim(sampling_matrix)[2], 2, replace = FALSE, prob = NULL)
  
  ## (Step 2) put sampled columns into "accumulative_matrix"
  accumulative_matrix <- cbind(accumulative_matrix, sampling_matrix[,random_2_cols])
  
  # (Step 3) sum 2 columns(species/clusters) picked and add it to the "sampling_matrix"
  sampling_matrix <- cbind(sampling_matrix, rowSums(sampling_matrix[,random_2_cols]))
  
  # (Step 4) remove the randomly picked 2 columns from the "sampling_matrix"
  sampling_matrix <- sampling_matrix[,-random_2_cols, drop = FALSE] # make the matrix not collapse into vector
  sampling_matrix
}

sampling_matrix
# include the [1,1,1,....,1] to the "accumulative_matrix"
accumulative_matrix <- cbind(accumulative_matrix,sampling_matrix)

# order the matrix
V <- accumulative_matrix
V <- V[,order(apply(V,2,sum), decreasing=T)] # sum of each column should decrease from left to right
# make the last n columns identity matrix
V <- cbind(V[,1:(n-1)],diag(n))

```

Then, get randomized branch length for 2n-1 branches and calculate the Σ = V x D(b) x VT.
```{r}
branch_num <- 2*n - 1
(B <- diag(runif(branch_num)))

# Phylogenetic closeness (distance of root to common ancestor)
Phylo_closeness = V %*% B %*% t(V)
Phylo_closeness

```

Scale so that we have an ultrametric tree.
- Distance of root to common ancestor (species 1 to 1, species 2 to 2) of the same species have to be equal (=1).\
- I couldn't figure this out. So I adopt Zach's method of randomizing internal branches' length and then adding the length of tips to consummate to 1.
```{r}
# Internal branches
B_int = diag(runif(n-1))
V_int = V[,-( (n) : (2*n-1) )]

Phylo_int = V_int %*% B_int %*% t(V_int)
Phylo_int
# the internal branch length is over 1, so we will make those less than 1 by reducing branch length.
B_int_half <- B_int * 0.2
Phylo_int = V_int %*% B_int_half %*% t(V_int)
Phylo_int

# Now as Zach's idea, use the same internal branch length + tip branch length (which is 1 - diagonal of Phylo_int)
scaled_branch_len <- c(diag(B_int_half),1-diag(Phylo_int))
Phylo_closeness_scaled = V %*% diag(scaled_branch_len) %*% t(V)
Phylo_closeness_scaled

```

Now solve the gLV model: dxdt=D(x)(1−Σx)
```{r}
# Find all feasible solutions
E <- Phylo_closeness_scaled
s <- rep(1,n)

# Stability <=> real part of eigenvalues are all negative
# function
leftmost <- function(M){
  eig <- eigen(M, only.values = T)$values
  # print(Re(eig)) # print all real part of eigenvalues
  return(min(Re(eig))) # the right most eigenvalue
} 

feasible_stable_list <- list()
for (subset in 1:(2^n-1)){
  presence_vector <- as.numeric(intToBits(subset)[1:n])
  s_subset <- s[presence_vector>0]
  E_subset <- E[presence_vector>0,presence_vector>0]
  x_star <- rep(0,n)
  x_star[presence_vector>0] <- solve(E_subset,s_subset)
  
  # check feasible & stable
  if((leftmost(E_subset) > 0) & (all(solve(E_subset, s_subset) > 0))){
    feasible_stable_list[[length(feasible_stable_list)+1]] <- list(
      community = subset,
      presence = presence_vector,
      xstar = x_star
    )
  }
}

# feasible_stable_list
length(feasible_stable_list) # all 31 community (2^5 -1 =31)
```
All feasible subset community are stable.

Lastly, make the assembly graph
```{r}
# adjacency matrix
adj_mat <- matrix(0,length(feasible_stable_list),length(feasible_stable_list))

# if community presence matrix of state i (e.g. 10010) is nested in state j (e.g. 10110), then we draw a directed link (i -> j)
# to test the nestedness, loop over all feasible, stable states' presence vector
for (i in 1:dim(adj_mat)[1]){
  for (j in 1:dim(adj_mat)[2]){
    # presence vectors
    vec_i <- feasible_stable_list[[i]]$presence
    vec_j <- feasible_stable_list[[j]]$presence
    if (all(vec_i <= vec_j)){ # i is nested to j
      # only 1 invasion event
      if (sum(vec_j - vec_i) == 1) {
        # print(paste0(vec_j - vec_i))
        adj_mat[i,j] <- 1
      }
    }
  }
}

library(igraph)
assembly_graph <- graph_from_adjacency_matrix(adj_mat)
plot(assembly_graph, layout = layout_with_sugiyama(assembly_graph)$layout)
```
The assembly graph is symmetric and organized (it shows all the combinations of nC1, nC2, nC3, nC4, nC5). It's because all of our feasible states are stable.



